---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
setwd("/Users/Tammy/online_courses/data_science/RepData_PeerAssessment1")
dat<- read.csv(unz("activity.zip","activity.csv"))
dat$date<- as.Date(dat$date)
```

## What is mean total number of steps taken per day?
```{r}
library(dplyr)
library(ggplot2)
dat<- tbl_df(dat)
steps_per_day<- dat %>% group_by(date) %>% summarise(steps_each_day=sum(steps, na.rm=T))
# make a histogram
ggplot(steps_per_day) + geom_histogram(aes(x=steps_each_day), color="white")
```

Calculate and report the mean and median of the total number of steps taken per day

```{r}
dat %>% group_by(date) %>% summarise(mean_steps=mean(steps, na.rm=T), 
                                     median_steps=median(steps, na.rm=T)) %>% print(n=61)
```


## What is the average daily activity pattern?
```{r}
interval_mean_steps<- dat %>% group_by(interval) %>% summarise(mean_steps_per_interval=mean(steps, na.rm=T))
ggplot(interval_mean_steps) + geom_line(aes(x=interval, y=mean_steps_per_interval))
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum  number of steps?

```{r}
interval_mean_steps %>% arrange(desc(mean_steps_per_interval)) %>% head(n=1)
```
## Imputing missing values
How many NAs in the orignal dataset?(how many rows contain NAs?)
```{r}
table(complete.cases(dat))
```
There are total 2304 rows containing NAs.

use the mean of that 5 min interval of all days to impute the missing NAs
```{r}
dat_no_NA<- dat
for (i in which(is.na(dat$steps))){
        idx<- interval_mean_steps$interval == as.numeric(dat[i,3])
        imputing_value<- as.numeric(interval_mean_steps[idx,2])
        dat_no_NA[i,1]<- imputing_value
}

```
histogram of total number of steps taken each day
```{r}
steps_per_day_no_NA<- dat_no_NA %>% group_by(date) %>% 
        summarise(steps_each_day=sum(steps, na.rm=T))
ggplot(steps_per_day_no_NA) + geom_histogram(aes(x=steps_each_day), color="white")
```

After imputing missing values, in some of the days, the mean steps are very close to the median 
steps
```{r}
dat_no_NA %>% group_by(date) %>% summarise(mean_steps=mean(steps, na.rm=T), 
                                     median_steps=median(steps, na.rm=T)) %>% print(n=61)
```

## Are there differences in activity patterns between weekdays and weekends?
add a column containing whether it is a weekday or weekend
```{r}
dat_weekday_weekend<- dat_no_NA %>% 
        mutate(weekday_or_weekend= ifelse(weekdays(date)=="Saturday" | 
                                                        weekdays(date) == "Sunday",
                                               "weekend", "weekday"))
dat_weekday_weekend$weekday_or_weekend<- as.factor(dat_weekday_weekend$weekday_or_weekend)
```
calculate the average steps every 5 min interval for weekdays and weekends
```{r}
weekday_weekend_mean_steps<- dat_weekday_weekend %>% group_by(weekday_or_weekend, interval) %>%
        summarise(mean_steps=mean(steps))
```

plot a line graph for the mean steps of weekdays and weekends respectively.

```{r}
ggplot(weekday_weekend_mean_steps) + geom_line(aes(x=interval, y=mean_steps,
                                                   colour=weekday_or_weekend))
```

or to make two plots in two panels

```{r}
ggplot(weekday_weekend_mean_steps) + geom_line(aes(x=interval, y=mean_steps)) +
        facet_grid(weekday_or_weekend ~ .)
```

### use facet_wrap rather than use facet_grid

```{r}
ggplot(weekday_weekend_mean_steps) + geom_line(aes(x=interval, y=mean_steps)) +
        facet_wrap(~weekday_or_weekend, ncol=1)
```

### it looks like this person took more steps over the weekends on average.